Xor-Some Exploratory Analyses for Preregistered Study: Model comparison SD grid search

### 2021-12-29 analysis/main/05_xor-some_analysis_model_compariso_grid-search-SD.Rmd
full SD grid search for all xor models vs Int only model

---
title: "Xor-Some Preregistered Study: Analyses"
author: "Polina Tsvilodub"
date: "8/31/2021"
output: github_document
---

@@

Since we observed potential colinearity effects in the main preregistered analysis for the trigger "or", we further explore the effects of this colinearity on the strength of the individual effects in the regression models.

To do so, we compute individual models omitting one of the factors on the "or" data, and compute the Bayes Factor (BF) of the model omitting one of the colinear predictors, relative to the full model.

@@

For exploration, we also compute the BF in favor of a model with an interaction between the two effects against the same model without interaction estimates.
```{r, results='hide', message=FALSE, warning=FALSE}
# use rather tight prior
priors <- set_prior("student_t(1, 0, 0.1)", class = "b")
# model with prior effect only
model_prior_comp_int <- brm(target ~ prior*competence +
                   (1 + prior + competence || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = 40000,
                 save_all_pars = TRUE
                 )
model_prior_comp <- brm(target ~ prior + competence +
                   (1 + prior + competence || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = 40000,
                 save_all_pars = TRUE
                 )
bf_pri_comp_int <- bayes_factor(model_prior_comp_int, model_prior_comp, silent = TRUE)
```

@@

>>>>

---
title: "Xor-Some Exploratory Analyses for Preregistered Study: Model comparison SD grid search"
author: "Polina Tsvilodub"
date: "12/1/2021"
output: github_document
---

@@

Since we observed potential collinearity effects in the main preregistered analysis for the trigger "or", we further explore the effects of this collinearity on the strength of the individual effects in the regression models.

To do so, we compute individual models omitting one of the factors on the "or" data, and compute the Bayes Factor (BF) of the model omitting one of the collinear predictors, relative to the full model.

@@

## Exploration of interaction effects

For exploration, we also compute the BF in favor of a model with an interaction between the two effects against the same model without interaction estimates. This is to estimate whether the additional complexity resulting from adding an additional estimate is benefitial in terms of model fit an hence beneficial as an exploratory factor.
```{r, results='hide', message=FALSE, warning=FALSE}
# use rather tight prior
priors <- set_prior("student_t(1, 0, 0.1)", class = "b")
# model with prior effect only
model_prior_comp_int <- brm(target ~ prior*competence +
                   (1 + prior + competence || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = 40000,
                 save_all_pars = TRUE
                 )
model_prior_comp <- brm(target ~ prior + competence +
                   (1 + prior + competence || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = 40000,
                 save_all_pars = TRUE
                 )
bf_pri_comp_int <- bayes_factor(model_prior_comp_int, model_prior_comp, silent = TRUE)
```

@@

## Model comparisons with grid search for final paper

Below, we conduct a similar grid search over more models. More specifically, we compute the BF in favor of a more complex model over the intercept-only baseline model. We compute these pairwise comparisons for each possible combination of factors over the intercept-only model, always adding the full interaction structure to the main effects. This results in 7 model comparisons. For each comparison, we perform a small grid search over the SD parameter of the prior t-distribution for the fixed effects regression coefficients, using the values [0.05, 0.1, 0.15]. For each model, 40.000 samples are extracted. These grid searches are performed separately for "or" and "some".
```{r}
# define a convenience function for iterating over the SDs and critical models
models_grid_search_SD_full = function(sd_vec, iter, trigger) {
  # create output df
  results <- tribble(
    ~"model in favor", ~"model against", ~"trigger", ~"SD", ~"BF", ~"log(BF)"
  )
  # filter data by trigger
  data <- d_critical_zScored_wide %>% filter(main_type == trigger)
  # models of interest
  formula_baseline <- "target ~ 1 + (1 | submission_id) + (1 | title)"
  formulas <- c(
    "target ~ prior + (1 + prior || submission_id) + (1 | title)",
    "target ~ relevance + (1 + relevance || submission_id) + (1 | title)",
    "target ~ competence + (1 + competence || submission_id) + (1 | title)",
    "target ~ relevance*competence + (1 + relevance + competence || submission_id) + (1 | title)",
    "target ~ prior*competence + (1 + prior + competence || submission_id) + (1 | title)",
    "target ~ prior*relevance + (1 + prior + relevance || submission_id) + (1 | title)",
    "target ~ prior*competence*relevance + (1 + prior + competence + relevance || submission_id) + (1 | title)"
  )
  model_name_list <- c("model_pri", "model_rel", "model_comp", "model_rel_comp", "model_pri_comp", "model_rel_pri", "model_pri_rel_comp")
  # set priors
  for (i in sd_vec) {
    dist <- paste("student_t(1, 0, ", i, ")", sep = "")
    prior <- set_prior(dist, class = "b")
    # fit the baseline intercept only model with same prior SD
    prior_int <- set_prior(dist, class = "Intercept")
    model_int <- brm(as.formula(formula_baseline),
                 data = data,
                 prior = prior_int,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = iter,
                 save_all_pars = TRUE,
                 silent = TRUE
                 )
    # fit all the models
    for (f in 1:length(formulas)) {
      print(formulas[f])
      model_target <- brm(as.formula(formulas[f]),
                 data = data,
                 prior = prior,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = iter,
                 save_all_pars = TRUE,
                 silent = TRUE
                 )
      # compute the BF
      print(paste("SD: ", i, sep=""))
      bf <- bayes_factor(model_target, model_int, silent = TRUE)
      print(paste("Model in favor: ", model_name_list[f], sep=""))
      print(paste("BF: ", bf, sep=""))
      # append to results dataframe, write out
      results <- results %>% add_row(
        "model in favor" = model_name_list[f],
        "model against" = "model_Int",
        "trigger" = trigger,
        SD = i,
        BF = round(bf[1]$bf, 4),
        "log(BF)" = round(log(bf[1]$bf), 4)
      )
    write_csv(results, 
              paste("../../data/main/grid_full_", trigger, "_SD_0_", i*100, ".csv", sep = ""))
    }

  }
  return (results)
}
```

```{r}
# call the grid search for xor 
grid_results_full_xor <- models_grid_search_SD_full(c(0.05, 0.1, 0.15), 40000, "xor")
# write out results
write_csv(grid_results_full_xor, "../../data/main/grid_full_xor_SD_final.csv")

# call the grid search for some 
grid_results_full_some <- models_grid_search_SD_full(c(0.05, 0.1, 0.15), 40000, "some")
# write out results
write_csv(grid_results_full_some, "../../data/main/grid_full_some_SD_final.csv")

Plot the effect of SD against the BF:
```{r}
grid_results_full_xor <- read_csv("../../data/main/grid_full_xor_SD_final.csv")
grid_results_full_some <- read_csv("../../data/main/grid_full_some_SD_final.csv")
grid_results_both <- rbind(grid_results_full_xor, grid_results_full_some)

grid_results_full_xor %>%
ggplot(., aes(x = SD, y = `log(BF)`, color = `model in favor`)) +
  geom_point() +
  geom_line() +
  facet_wrap(~trigger) + 
  ggtitle("BF in favor of including certain effects over an Intercept-only model,\ndepending on SD of the prior t-distribution over betas")

ggsave("BF_vs_SD_allModels_vs_int.pdf", width = 6.5, height = 4)
```

## Additional interaction comparisons

Compute the critical BF for including the effect of relevance over including the effect of prior in a two-factor model when only including the simple effects, no interactions.
```{r}
prior <- set_prior("student_t(1, 0, 0.05)", class = "b")
model_rel_comp_simple <- brm(target ~ competence+relevance +
                   (1 + competence + relevance || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide %>% filter(main_type == "xor"),
                 prior = prior,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = 40000,
                 save_all_pars = TRUE,
                 silent=TRUE
                 )

model_pri_comp_simple <- brm(target ~ competence+prior +
                   (1 + competence + prior || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide %>% filter(main_type == "xor"),
                 prior = prior,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = 40000,
                 save_all_pars = TRUE
                 )
bf_rel_pri_simple <- bayes_factor(model_rel_comp_simple, model_pri_comp_simple, silent=TRUE) 
# read old data with interaction comparisons
df_ints <- read_csv("./../../data/main/bf_interactions_comparisons.csv")
glimpse(df_ints)
df_ints %>% add_row(
  `Model in favor` = "Rel + Comp",
  `Model against` = "Pri + Comp",
  BF = round(bf_rel_pri_simple[1]$bf, 4),
  SD = 0.05
) #%>% write_csv("./../../data/main/bf_interactions_comparisons.csv")
```

```{r}
# TODO: use different seeds / rerun the process for a more stable BF estimate
```
